
%
%  $Description: Author guidelines and sample document in LaTeX 2.09$ 
%
%  $Author: ienne $
%  $Date: 1995/09/15 15:20:59 $
%  $Revision: 1.4 $
%

\documentclass[times, 10pt,twocolumn]{IEEEtran} 
\usepackage{latex8}
\usepackage{times}




%\documentstyle[times,art10,twocolumn,latex8]{article}

%------------------------------------------------------------------------- 
% take the % away on next line to produce the final camera-ready version 
\pagestyle{empty}

%------------------------------------------------------------------------- 
\begin{document}

\title{Title of the document here}

% author names and affiliations
\author{\IEEEauthorblockN{
  John Burchell\IEEEauthorrefmark{1} and
  William Granli\IEEEauthorrefmark{1}}

  \IEEEauthorblockA{\IEEEauthorrefmark{1}john.a.burchell, william.granli@gmail.com}

}
\maketitle
\thispagestyle{empty}

\begin{abstract}
   The ABSTRACT is to be in fully-justified italicized text, at the top 
   of the left-hand column, below the author and affiliation 
   information. Use the word ``Abstract'' as the title, in 12-point 
   Times, boldface type, centered relative to the column, initially 
   capitalized. The abstract is to be in 10-point, single-spaced type. 
   The abstract may be up to 3 inches (7.62 cm) long. Leave two blank 
   lines after the Abstract, then begin the main text. 
\end{abstract}



%------------------------------------------------------------------------- 
\Section{Introduction}
%Problem and definition
Comparisons between programming languages are difficult to perform, due to inherent differences between languages, such as syntax. One aspect of code is its readability, which in turn is a determining factor of a piece of software's quality and maintainability \cite{aggarwal2002integrated, elshoff1982improving}. Buse \& Westley define readability as ``a property that influences how easily a given piece of code can be read and understood'' \cite{buse2010learning}. 
\newline

%%Existing literature
Existing studies in the field focus on readability within of different coding styles for a set language, which can be seen in for example \cite{buse2010learning}. Buse \& Westley introduce a metrics system for measuring the readability of code, and Java is the language of choice in the study \cite{buse2010learning}. Buse \& Westley also discuss readability and its impact on programming language design, but they do not explore the readability differences between languages \cite{buse2010learning}. Posnett et al. \cite{posnett2011simpler} have further developed Buse \& Westley's metrics model to provide a simpler and more practical model which builds on size metrics as well as Halstead metrics \cite{halstead1977elements}. Hanen et al. have presented a study in the field which aims to identify details in code which makes it hard to read and the main focus is put on syntactical elements such as indentation or curly brackets \cite{hansen2013makes}. Even though direct comparisons between languages were not presented in the study \cite{hansen2013makes}, conclusions related to this can be drawn thanks to the fact that these syntactical elements only exist in some languages. 
\newline

The importance of code readability may be overlooked, since it does not directly affect runtime performance. Aggarwal et al., however claim that code's readability impacts it's maintainability \cite{aggarwal2002integrated}. This claim is also supported by \cite{elshoff1982improving} et al. who also introduce examples of certain coding practices that are more readable and maintainable than others \cite{elshoff1982improving}.


%------------------------------------------------------------------------- 
\Section{Purpose of the Study}

%Introduction, specifying why this is urgent, needed etc from the notes above

Industry and academia both face the problem of code maintainability, and by association, readability and as was highlighte previously, the readability of code can affect its maintainability. This study could help some in industry to decide which languages should be used if maintainability is of high concern. Further studies could be conducted by to determine if there is a correlation between the popularity of projects and the readability of their languages. Furthermore, academia would be interested in which languages are more popular and those that have a higher readability, especially for novice developers. Further investigation into the design of these programming languages could yield insights into what makes them so popular and potentially well suited for teaching.
\newline

The purpose of this survey study is therefore to test what are the differences in readability between five common and well known programming languages. With the aim of suggesting a metric which describes the readability of a given program, taking into account the inherent readability of the language in which it is written.
\newline

The purpose of this survey study is therefore to find what are the differences in readability between five common and well known programming languages. With the aim of developing a metric to describe the readability of a given program, which takes into account the inherent readability of the language in which it is written.
\newline

The Independent variables will be defined as the experience, occupation, preferred language, least preferred language and the working language of the participant. The dependent variable in this example will be the readability of the language.
\newline

The survey data will be collected from professionals, teachers, students and hobbyists via a series of on-line surveys. The respondents will then be asked some information about their development background. The will then be asked to review a series of code snippets from popular languages and to rate their readability. 
\newline

%------------------------------------------------------------------------- 
\SubSection{Research Question}
This study has been designed with one main research question in mind:
\begin{itemize}
\item \textit{RQ1:} What are the measurable differences in readability between programming languages?
\end{itemize}


The results of RQ1 will then be used to answer the following two sub research questions. 
\begin{itemize}
\item \textit{SRQ1}: How do the measurable differences between languages correlate to their use in academia for teaching?
\item \textit{SRQ2}: How do the measurable differences between languages correlate to the popularity of the language?
\end{itemize}


%------------------------------------------------------------------------- 
\Section{Methodology}
The main purpose of this study is to discover differences in the perceived readability of code. The study will be carried out in three phases. In phase 1, we will identify the target population. In phase 2, the survey will be designed and carried out. In phase 3, the results of the survey will be analysed. The investigation method will be in the form of a questionnaire. The main argument for the questionnaire is that the survey must be distributed to a large amount of people. A quantitative study is also well-suited for this study since there are no existing publications that explore readability differences between languages. 

Longitudinal servey - Collected over time (cross-sectional is one isntance in time)
Purpose: To generalise perceived readability of different programming languages by novices - Ref papers from the book
Indicate why a survey is the preferred type of data collection - Consider advantages of surveys (economy and design and rapid data collection - (Fowler 2009) -> discuss advantage of identifying attributes of a large population from a small group of individuals 
Form: Internet based (nesbary 2000, sue \& ritter 2012) - Provide a rationale for its use, strengths and weaknesses. Costs, data availability and convenience
  peer reviewed with collegues, tested on a small group that we supervise

  Population: limited to the SEM program. (Count students if it's available somewhere).
  Cluster sampling / single stage - Not possible to get all the names of all possible respondents - Babbie 2007
  nonprobability sample - convenience sample - chosen on their availability and convenience. Not ideal, but best we can probably do considering its single stage etc (ideally we would do random sampling)
  stratification will not be enforced. Ideally we will have 1/3,1/3,1/3 but we cannot enfore. The groups are open and public
  total at SEM = ~300. (Count how many in the facebook groups, that's the sample size - Estimate how many will actually respond, how much error we can tolerate (\%4 probably - confidence interval) - Use the sample size method by fowler?


\SubSection{Identifying the Population}
As mentioned before, the goal of this study is to identify the inherent differences in readability of programming languages. To research this, a broad population is required, to avoid results influenced by a possible correlation between the perception of readability and the previously defined independent variables. As such, a research survey has been conducted. This reasoning is reinforced by Easterbrook et al. \cite{easterbrook2008selecting} who state ``Survey research is used to identify the characteristics of a broad population of individuals''. 
\newline

The population of this survey is the students of the Software Engineering and Management programme (SEM) at Gothenburg University. The industry is left out of the population since they are more likely to have a lot of experience with certain languages which may cause them to be biased towards that language. Furthermore, Höst, Regnell and Whohlin \cite{host2000using} make the case that ``The reason to use students as subjects is often that they are available at universities
and they are willing to participate in studies as part of courses they attend.'' which is highly approprpiate given that we will be asking our peers and those in lower years who will eventually also be required to perform research tasks. The SEM program was chosen due to it being accessible for the researchers. 

99 - 3rd year
168 - 2nd year
190 - 1st year

Overlap: 50 (Those who are in another group)
Total: 397

Population in the facebook groups.

count here - estimate desired response rate, error rate etc
Sampling bias is a considerable challenge in research \cite{easterbrook2008selecting}

\newline

\SubSection{Questionnaire Design}

The five languages represented in the questionnaire were C++, Erlang, Go, Java and Python. They were chosen for multiple reasons. One being that all the students in the SE\&M course will have taken an introductory Java programming course by now and as such should be familiar with the syntax. The other languages were chosen for similar reasons. Erlang was chosen as it's the first language taught after Java and it uses a different development paradigm. C++ was chosen as it is used heavily in project course during the 4th semester and python was chosen for similars reasons. Golang was picked as we wanted to see how students would perceive a language that is popular but has not really entered the academic landscape as of yet.

Google Sheets was decided upon as the appropriate tool of choice for our survey. It has a good reputation for use in surveys and it was recommended to us by Branson. We also have prior experience the tool from previous research projects.
The questionnaire consists of structured and partially structured questions. The questions are followed by two code snippets per language mentioned previously with a scale requesting the respondent to rate whether they feel that the code is easy or difficult to read in the style of the Likert Scale. The questions were written with as little ambiguity as possible. They were field tested by colleagues in a small example of our survey. Examples of these questions can be found in ***Appendix Item Here***. The aforementioned chart shows how some of our independent variables, research questions and survey items correlate.

Taking into consideration potential biases that may occur, the code snippets were designed so that the code should be easily understood by first year students who have only had exposure to Java. For example, each code snippet has a standard ``Hello World'' example and a simple for loop with a counter that is returned at the conclusion of the loop. The only exception to this is Erlang. While erlang does support for loops, recursion is the tool of choice to perform iterations. As such, we feel it was a valid example and should not affect the results for this specific language drastically considering, by this point and time in their education, all SE\&M students will have been exposed to recursion. The choices of languages do leave us open to a potential validity and bias issue. Inexperienced students who have only seen Java may rate it as easily readable purely because they are only familiar with Java, this is why the relative level of experience is requested first in the survey, so that we might be able to pick up this potential bias during the results analysis. To help alleviate this issue, the questionnaire was distributed to all three year groups currently studying at SE\&M (2012,2013,2014).

Other potential validity threats we must take into consideration are those that arise from the nature of survey. Since we are performing single stage data collection, the respondents of the surveys are answering our questions at their convenience. This opens the door for some validity threats, such as the willingness to participate. Our data are at the mercy of those willing to participate; we cannot anticipate how many will respond. There is little incentive for the respondents to partake in the survey other than for their own self interest.

\SubSection{Results Analysis}
The response rate for the questionnaire was 11\%. The questionnaire was distributed to 270 members of the sample group and 29 members returned the survey. An additional 6 students have answered the questionnaire in a second iteration, where they were personally contacted based on which year they were attending. Table XXX1 displays information of the respondents experience level. Table XXX2 shows the year which the students are attending. 
\newline

Response bias has been determined mostly through analysing the correlation between the independent variables and the dependent variables. Readability of a language is likely to be strongly correlated with the reader's experience of the language. Due to there being a high response rate amongst 1st year students, it suggests that the nonrespones may have affected the results. This is not necessarily negative, since our study aims to measure the inherent readability of a programming language, where the experience of the reader has as little effect as possible. 
\newline

The data was collected with a questionnaire consisting of 15 questions. The data has been analysed with simple descriptive analysis where mean values have been the main source to analyse the readability of a language (RQ1). Correlation between the independent and dependent variables have been the main analysis to answer SRQ1 and SRQ2. Score ranges have been used to analyse the strength of replies. More advanced data analysis methods have not been utilized due to the relatively low number of respondants. 

\SubSection{Threats to Validity}
As in all research there are threats to the validty of the study being undertaken. Our largest threat would be the external validty of the data collected. The students were selected from the three current years studying at the University of Gothenburg. Most of the respondants indicated that they are in the first year of their studties and as such their inexperience could be a limiting factor on their answers to the survey. The students may also have had a bias for one of the languages; Java. So far, the students in the first year have been formally taught in only Java. This means that there is a lower chance they would be familiar with reading other language syntax other than java and as such they would be more likely to see Java favourably than others. We cannot therefore generalise that Java has higher readability than that of the other languages despite what the results show. 

A further threat that we were unable to counteract adequately was the willing to participate of our respondants. The participants had little or no incentive to actually partake in the study and as such many most likely simply ingored it. This is to be expected when one attempts to undertakes a single stage sampling method. One idea to improve the response rate for further surveys could be to form some incentive for the respondants.

The fact that the questionnaire was distributed online via facebook groups also meant that there was another threat to the validty of the responses. These groups are not offically controled by the university and as such admission into these groups is up to the students who are organising them. This means that there is potential that some of the responses came from people not within the SE\&M isntitutions and this should be taken into consideration when analysing the results.

\bibliographystyle{assignment_1}
\bibliography{assignment_1}

\begin{table}[ht]
  \centering
  \begin{tabular}{| l | r | c |}
  	\hline
  	\textbf{Variable Name} & \textbf{Research Question} & \textbf{Item on Survey} \\ \hline
  	Experience & RQ1, SQ1 & 1 \\ \hline
    Academic Year & RQ1, SQ1 & 2 \\ \hline
  	Preferred Language & RQ1, SQ2 & 3 \\ \hline
  	Least Preferred Language & RQ1, SQ2& 4 \\  \hline
  	Experience of Language & RQ1, SQ2 & 5 \\ 
  	\hline
	\end{tabular}
  \caption{ Hiadsdsafdsaf}
\end{table}

\end{document}

